---
title: "NBLAST"
---

```{r setup, include=FALSE}
source('loadNatPackages.R')
loadNatPackages()
source('nat-startup.R')
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
knit_hooks$set(webgl = hook_webgl)
```

```{r global_options,include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
rotationMatrix=rbind(c(1,0,0,0),c(0,-1,0,0),c(0,0,-1,0),c(0,0,0,1))
windowRect=c(1221,231,1767,700)
```

The package [nat.nblast](http://natverse.org/nat.nblast/) implements the NBLAST neuron similarity algorithm described in [Costa et al. 2016](http://doi.org/10.1016/j.neuron.2016.06.012). 

<font size=5> Using NBLAST to assess neuron similarity </font>

NBLAST can be used to assess the similarity between neurons. In this example, NBLAST is used to compare a neuron drawn from a light-level tracing and a neuron drawn from an EM reconstruction (plotted in black and blue, respectively).
```{r, webgl=TRUE, echo=TRUE,results='hide',message=FALSE,warning=FALSE}
neuronTracing <- read.neuron('data/registeredNeuron_JFRC2013.swc',class="neuron")
A2_EM<- read.neuron('data/DNa02_01_09_2020.swc')
neuronEM<- xform_brain(A2_EM,sample=FAFB, reference = JFRC2013)

#plot both neurons 
open3d(userMatrix=rotationMatrix,windowRect=windowRect,zoom=0.6)
plot3d(neuronTracing,lwd=3,col='black',WithNodes=FALSE)
plot3d(neuronEM,lwd=3,col='blue',WithNodes=FALSE)
plot3d(JFRC2013)
```
```{r,include=FALSE}
clear3d()
```
```{r,echo=TRUE,results='hide',warning=FALSE,message=FALSE}
all_neurons=neuronlist(neuronTracing,neuronEM)
nblast_scores=nblast_allbyall(all_neurons,smat = NULL, distance = FALSE)
```

```{r,echo=FALSE}
neuron_names=c('neuronTracing','neuronEM')
scores<- matrix(ncol=length(all_neurons),nrow=length(all_neurons))
for (i in 1:length(all_neurons)){
    for (j in 1:length(all_neurons)){
      scores[i,j]=nblast(all_neurons[i],all_neurons[j],normalised=TRUE)
    }
}
normalized_scores=data.frame(scores)
rownames(normalized_scores) <- neuron_names
colnames(normalized_scores)<-neuron_names
normalized_scores
```

<font size=5> Using NBLAST to find flycircuit clones </font>

NBLAST can also be used to query databases of neurons based on similarity to your neuron of interest. 

First, load in your neuron of interest 
```{r,webgl=TRUE,echo=TRUE,results='hide',message=FALSE}
g13 <- read.neuron('data/g13_JFRC2013.swc',class="neuron")
open3d(userMatrix=rotationMatrix,windowRect=windowRect,zoom=0.6)
plot3d(g13,lwd=3,col='black',WithNodes=FALSE)
g13_JFRC2=xform_brain(g13,sample=JFRC2013,reference=JFRC2)
```

To find flycircuit clones,download the flycircuit data and convert it to a dotprops object(this takes a while but only needs to be done once )
```{r,include=FALSE}
clear3d()
```
```{r A2plots, echo=TRUE,results='hide',message=FALSE,warning=FALSE}
fc_download_data('http://flybrain.mrc-lmb.cam.ac.uk/si/nblast/flycircuit/allbyallblastcv4.5.ff',
                 type='ff')
options('flycircuit.scoremat'="allbyallblastcv4.5.ff")
dps<-read.neuronlistfh("http://flybrain.mrc-lmb.cam.ac.uk/si/nblast/flycircuit/dpscanon.rds",
                       localdir=getOption('flycircuit.datadir'))
options('nat.default.neuronlist'='dps')
if(!require("devtools")) install.packages("devtools")
devtools::source_gist("bbaf5d53353b3944c090", filename = "FlyCircuitStartupNat.R")
dpscanon=read.neurons(dps)
```
Next, register your neuron to the FCWB template brain (the template brain space to which the flycircuit neurons are registered).
If you want to search for bilateral hits, mirror your neuron about the y-axis of the FCWB template brain
Convert this registered neuron to a  [dotprops](http://natverse.org/nat/reference/dotprops.html) object.
```{r, webgl=TRUE, echo=TRUE,results='hide',message=FALSE}
#transform brain into FCWB template space 
g13_FCWB=xform_brain(g13_JFRC2,sample=JFRC2,reference=FCWB)
g13_FCWB_mirror=mirror_brain(g13_FCWB,brain=FCWB)

#plot the neuron and its mirrored version 
open3d(userMatrix=rotationMatrix,windowRect=windowRect,zoom=0.6)
plot3d(g13_FCWB,lwd=3,col='black',WithNodes=FALSE)
plot3d(g13_FCWB_mirror,lwd=3,col='black',WithNodes=FALSE)

#convert neuron to dotprops object 
g13_dotprops=dotprops(c(g13_FCWB,g13_FCWB_mirror))

```

Finally, call nbast with your neuron and the flycircuit data as inputs 
```{r,include=FALSE}
clear3d()
```
```{r,echo=TRUE}
nblast_fc=nblast(g13_dotprops,target = dpscanon, normalised = TRUE);

#display top 10 hits 
nblast_fc_df=data.frame(name=names(nblast_fc),score=as.numeric(nblast_fc))
nblast_fc_df[order(-nblast_fc_df$score),][1:10,]
```

<font size=5> Using NBLAST to find Gal4 lines</font>

To find Gal4 lines similar to your neuron of interest, first download the Gal4 data. This takes a while but only needs to be done once. (Note: this database contains only GMR lines)
```{r,echo=TRUE}
gmrdps<-read.neuronlistfh("http://flybrain.mrc-lmb.cam.ac.uk/si/nblast/gmrdps/gmrdps.rds",
                          localdir=getOption('flycircuit.datadir'))
```

Next, register your neuron to the FCWB template brain (the template brain space to which the Gal4 line data is registered).
If you want to search for bilateral hits, mirror your neuron about the y-axis of the FCWB template brain
Convert this registered neuron to a  [dotprops](http://natverse.org/nat/reference/dotprops.html) object.

```{r, webgl=TRUE, echo=TRUE,results='hide',message=FALSE}
#transform brain into FCWB template space 
g13_FCWB=xform_brain(g13_JFRC2,sample=JFRC2,reference=FCWB)
g13_FCWB_mirror=mirror_brain(g13_FCWB,brain=FCWB)

#plot the neuron and its mirrored version 
open3d(userMatrix=rotationMatrix,windowRect=windowRect,zoom=0.6)
plot3d(g13_FCWB,lwd=3,col='black',WithNodes=FALSE)
plot3d(g13_FCWB_mirror,lwd=3,col='black',WithNodes=FALSE)

#convert neuron to dotprops object 
g13_dotprops=dotprops(c(g13_FCWB,g13_FCWB_mirror))
```

Finally, call nbast with your neuron and the Gal4 data as inputs
```{r,include=FALSE}
clear3d()
```
```{r,echo=TRUE}
nblast_gal4=nblast(g13_dotprops,target = gmrdps, normalised = TRUE);

#display 10  hits
nblast_gal4_df=data.frame(name=names(nblast_gal4),score=as.numeric(nblast_gal4))
nblast_gal4_df[order(-nblast_gal4_df$score),][1:10,]
```
