---
title: " Working with Hemibrain Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(rgl)
knit_hooks$set(webgl = hook_webgl)
source('nat-startup.R')
```

```{r global_options,include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

Data from the Janelia hemibrain release is available at: https://neuprint.janelia.org/?dataset=hemibrain:v1.0.1&qt=findneurons 
The package [neuprintr](http://natverse.org/neuprintr/reference/index.html) can be used to work with this data in R 

<font size=5> Fetch Hemibrain Skeletons </font>
To access hemibrain data using R, you first have to login to the neuPrint server using the [neuprint_login](http://natverse.org/neuprintr/reference/neuprint_login.html) function. Find your token at https://neuprint.janelia.org/account\

```{r,echo=TRUE,results='hide',message=FALSE,warning=FALSE,eval=FALSE}
conn = neuprint_login(server= "https://neuprint.janelia.org/",
                      token= "yourtokenhere")
```
```{r, include=FALSE}
conn = neuprint_login(server= "https://neuprint.janelia.org/",
                      token= "eyJhbGciOiJIUzI1NiIsInR5cCI6IkpXVCJ9.eyJlbWFpbCI6ImlzYWJlbGFkYWxlc3NhbmRyb0BnbWFpbC5jb20iLCJsZXZlbCI6Im5vYXV0aCIsImltYWdlLXVybCI6Imh0dHBzOi8vbGgzLmdvb2dsZXVzZXJjb250ZW50LmNvbS9hLS9BT2gxNEdqRVMzRG9INXBPUTU4Rmh6TUQwQnlyYUg0RFNPSUpBMzA3NVoycT9zej01MD9zej01MCIsImV4cCI6MTc2NzI0NzM3MH0.WLZrEFi2rK9E6z5RCU_NHToY1gI7O2t6uLnJ5nGt5vg")
hemibrainMatrix=rbind(c(1,-0.03,-0.04,0),c(-0.04,0,-1,0),c(0.04,1,0,0),c(0,0,0,1))
rotationMatrix=rbind(c(1,0,0,0),c(0,-1,0,0),c(0,0,-1,0),c(0,0,0,1))
windowRect=c(1221,231,1767,700)
```
There are many ways to find a neuron, including by name, bodyid, or ROI innervation. Here, we find all hemibrain neurons with names matching 'R1'
```{r, webgl=TRUE, echo=TRUE}
R1.info = neuprint_search("R1_.*")

#read in the skeletons as neurons 
R1_neurons = neuprint_read_neurons(R1.info$bodyid)
open3d(userMatrix=hemibrainMatrix,windowRect=windowRect,zoom=0.7)
plot3d(R1_neurons,lwd=2,WithNodes=FALSE,soma=TRUE)
```
<font size=5> Fetch brain regions </font>
You can see a list of all available brain region meshes using the following command(*Note: not all of these meshes are currently available for download):
```{r,include=FALSE}
clear3d()
```
```{r}
neuprint_ROIs()
```
To fetch these messages, use [neuprint_ROI_mesh](http://natverse.org/neuprintr/reference/neuprint_ROI_mesh.html): 
```{r, webgl=TRUE, echo=TRUE}
#plot our neurons of interest 
open3d(userMatrix=hemibrainMatrix,windowRect=windowRect,zoom=0.7)
plot3d(R1_neurons,lwd=2,WithNodes=FALSE,soma=TRUE)

#add EB and LAL meshes 
EB=neuprint_ROI_mesh(roi = "EB")
plot3d(EB, add = TRUE, alpha = 0.1)
LAL_R=neuprint_ROI_mesh(roi = "LAL(R)")
plot3d(LAL_R, add = TRUE, alpha = 0.1)
```
```{r,include=FALSE}
clear3d()
```
<font size=5> Compare with neurons in other brain spaces</font>
The JRCFIB2018F template brain can be used to transform skeletons fetched from the hemibrain into other tempate brain spaces for comparisons with other neurons. To download the template brain and necessary registration files, run the following code (this takes a while but only needs to be done once).
```{r,echo=TRUE,results='hide',message=FALSE,warning=FALSE,eval=FALSE}
remotes::install_github("natverse/nat.jrcbrains")
nat.jrcbrains::download_saalfeldlab_registrations()
```
Then apply the transformations. In this example, the hemibrain data is being transformed into the FCWB template brain space so that it can be used to find similar flycircuit neurons or gal4 lines via nblast. 
*Note: [Java](https://www.oracle.com/java/technologies/javase-downloads.html) and [rJava](https://github.com/s-u/rJava) need to be installed in order to run the transformation: 
```{r, webgl=TRUE, echo=TRUE,results='hide',message=FALSE,warning=FALSE}
#first transform into an intermediate brain space for which the transofrmation is defined 
R1.jrc2018F = xform_brain(R1_neurons*8/1000, reference="JRC2018F", sample="JRCFIB2018F")

#then transform into FCWB brain space 
R1.fcwb = xform_brain(R1.jrc2018F, reference='FCWB', sample='JRC2018F')

#plot neurons with fcwb template brain 
open3d(userMatrix=rotationMatrix,windowRect=windowRect,zoom=0.6)
plot3d(R1.fcwb,lwd=2,WithNodes=FALSE)
plot3d(FCWB)
```

<font size=5> Find synapse locations</font>
```{r, webgl=TRUE, echo=TRUE,results='hide',message=FALSE,warning=FALSE}
R1_synapses=neuprint_get_synapses(R1.info$bodyid)

open3d(userMatrix=hemibrainMatrix,windowRect=windowRect,zoom=0.7)
plot3d(R1_neurons,col='black',lwd=2,WithNodes=FALSE,soma=TRUE)
points3d(xyzmatrix(R1_synapses),col='red',lwd=2)
```
```{r,include=FALSE}
clear3d()
```
<font size=5> Get synaptic partners</font>
```{r,echo=FALSE,results='hide',message=FALSE,warning=FALSE}
R1_all_inputs.neurons <- read.neurons('data/R1_all_input_neurons.swc',class="neuron")
R1_all_outputs.neurons <- read.neurons('data/R1_all_output_neurons.swc',class="neuron") 
```
Finding all the <span style="color:red"> presynaptic partners of the R1 neurons 
```{r,echo=TRUE,results='hide',message=FALSE,warning=FALSE}
#get all inputs to R1 neurons 
R1_all_inputs=unique(neuprint_connection_table(R1.info$bodyid,prepost='PRE')$partner)
```
```{r,echo=TRUE,results='hide',message=FALSE,warning=FALSE,eval=FALSE}
R1_all_inputs.neurons=neuprint_read_neurons(R1_all_inputs$bodyid)
```
```{r, webgl=TRUE, echo=TRUE,results='hide',message=FALSE,warning=FALSE}
open3d(userMatrix=hemibrainMatrix,windowRect=windowRect,zoom=0.7)
plot3d(R1_neurons,col='black',lwd=2,WithNodes=FALSE,soma=TRUE)
plot3d(R1_all_inputs.neurons,col='red',lwd=2,WithNodes=FALSE,soma=TRUE)
```
```{r,include=FALSE}
clear3d()
```
Finding all the <span style="color:blue"> postsyanptic partners of the R1 neurons 
```{r,echo=TRUE,results='hide',message=FALSE,warning=FALSE}
R1_all_outputs=unique(neuprint_connection_table(R1.info$bodyid,prepost='POST')$partner)
```
```{r,echo=TRUE,results='hide',message=FALSE,warning=FALSE,eval=FALSE}
R1_all_outputs.neurons=neuprint_read_neurons(R1_all_outputs)
```
```{r, webgl=TRUE, echo=TRUE,results='hide',message=FALSE,warning=FALSE}
open3d(userMatrix=hemibrainMatrix,windowRect=windowRect,zoom=0.7)
plot3d(R1_neurons,col='black',lwd=2,WithNodes=FALSE,soma=TRUE)
plot3d(R1_all_outputs.neurons,col='blue',lwd=2,WithNodes=FALSE,soma=TRUE)

```
Next,group these inptus and outputs by type 
```{r, webgl=TRUE, echo=TRUE,results='hide',message=FALSE,warning=FALSE}
library(plyr)
#group the inputs by type 
R1_all_inputs_count=count(neuprint_get_meta(R1_all_inputs),'type')
R1_all_inputs_count=subset(R1_all_inputs_count,freq>5)
R1_all_inputs_count=subset(R1_all_inputs_count,type!='NA')

#group the outputs by type 
R1_all_outputs_count=count(neuprint_get_meta(R1_all_outputs),'type')
R1_all_outputs_count=subset(R1_all_outputs_count,freq>5)
R1_all_outputs_count=subset(R1_all_outputs_count,type!='NA')

#aggregate data and plot as a bar graph 
R1_all_partners=merge(R1_all_inputs_count,R1_all_outputs_count,by='type',all=TRUE)
R1_all_partners.m=melt(R1_all_partners, id.vars='type')
R1_all_partners.m=R1_all_partners.m[order(-R1_all_partners.m$value),]
R1_all_partners.m$Groups <- factor(R1_all_partners.m$type, levels = R1_all_partners[order(R1_all_partners$freq.x),]$type)
plot<-ggplot(R1_all_partners.m,aes(fill=variable,x=Groups,y=value))+geom_bar(stat="identity")+
  coord_flip()+theme_bw()+scale_x_discrete(name ="Region")+ ylab('Synapse Count')+theme(legend.position="none")+scale_fill_manual(values=c("blue","red"))
print(plot)
```

We can also look only at the synaptic partners which are common to the set of all R1 neurons using neuprint_common_connectivity
```{r,echo=FALSE,results='hide',message=FALSE,warning=FALSE}
R1_common_inputs=neuprint_common_connectivity(R1.info$bodyid,prepost='PRE')
R1_common_inputs.neurons=read.neurons('data/R1_common_input_neurons.swc',class="neuron")

R1_common_outputs=neuprint_common_connectivity(R1.info$bodyid,prepost='POST')
R1_common_outputs.neurons=read.neurons('data/R1_common_output_neurons.swc',class="neuron")
```
```{r,echo=TRUE,results='hide',message=FALSE,warning=FALSE,eval=FALSE}
R1_common_inputs=neuprint_common_connectivity(R1.info$bodyid,prepost='PRE')
R1_common_inputs.neurons=neuprint_read_neurons(as.character(colnames(R1_common_inputs)))

R1_common_outputs=neuprint_common_connectivity(R1.info$bodyid,prepost='POST')
R1_common_outputs.neurons=neuprint_read_neurons(as.character(colnames(R1_common_outputs)))
```
```{r, webgl=TRUE, echo=TRUE,results='hide',message=FALSE,warning=FALSE}
#plot common inputs and output 
open3d(userMatrix=hemibrainMatrix,windowRect=windowRect,zoom=0.7)
plot3d(R1_neurons,col='black',lwd=2,WithNodes=FALSE,soma=TRUE)
plot3d(R1_common_inputs.neurons,col='red',lwd=2,WithNodes=FALSE,soma=TRUE)
plot3d(R1_common_outputs.neurons,col='blue',lwd=2,WithNodes=FALSE,soma=TRUE)
```
```{r,include=FALSE}
clear3d()
```
Grouping these inputs and outputs by type, we get: 
```{r, webgl=TRUE, echo=FALSE,results='hide',message=FALSE,warning=FALSE}
R1_common_inputs_df=neuprint_get_meta(as.character(colnames(R1_common_inputs)))

R1_common_outputs_df=neuprint_get_meta(as.character(colnames(R1_common_outputs)))

R1_input_type=count(R1_common_inputs_df,'type')
R1_output_type=count(R1_common_outputs_df,'type')
R1_common_partners=merge(R1_input_type,R1_output_type,by='type',all=TRUE)
R1_common_partners.m=melt(R1_common_partners, id.vars='type')
R1_common_partners.m=R1_common_partners.m[order(-R1_common_partners.m$value),]
R1_common_partners.m$Groups <- factor(R1_common_partners.m$type, levels = R1_common_partners[order(R1_common_partners$freq.y),]$type)
R1_plot=ggplot(R1_common_partners.m,aes(fill=variable,x=Groups,y=value))+geom_bar(stat="identity")+ ggtitle('R1a common partners')+
  coord_flip()+theme_bw()+scale_x_discrete(name ="Neuron Type")+ ylab('# Connections')+theme(legend.position="none")+scale_fill_manual(values=c("blue","red"))
print(R1_plot)
```

<font size=5> Evaluate neuron connectivity</font>
The neuprint_get_adjacency_matrix function can be used to generate an adjacency matrix between two sets of neurons. Here, we look at the connectivity between R1 and EPG neurons. 
```{r, webgl=TRUE, echo=TRUE,results='hide',message=FALSE,warning=FALSE}

#find all EPG neurons 
EPG.info=neuprint_search(".*EPG.*")

#get adjacency matrix 
adj_mat=neuprint_get_adjacency_matrix(c(R1.info$bodyid,EPG.info$bodyid))

#plot results as a heatmap
rownames(adj_mat)=vector(mode='character',length=dim(adj_mat)[1])
colnames(adj_mat)=vector(mode='character',length=dim(adj_mat)[2])
Heatmap(adj_mat,row_title='R',column_title='EPG',show_row_dend=FALSE,
        show_column_dend=FALSE,show_row_names=FALSE,show_column_names=FALSE,heatmap_legend_param=list(title='weight'))

```

<font size=5> Get the shortest paths </font>